---
title: "developmental_brainage_analysis"
output: html_document
---

load libraries

```{r}
library(ggplot2)
library(readr)
library(readxl)
```

load files

```{r}

#Male data
MR_predicted_age_male <- read_csv("/KIMEL/tigrlab/scratch/jshah/brain_age_analysis/data/processed/developmental_brainage_output/output_file_2024-10-22-22-50-10_MR_predicted_age_male.csv", show_col_types = FALSE)

BrainAGE_male <- read_csv("/KIMEL/tigrlab/scratch/jshah/brain_age_analysis/data/processed/developmental_brainage_output/output_file_2024-10-22-22-50-10_BrainAGE_male.csv", show_col_types = FALSE)

Adjusted_MR_predicted_age_male <- read_csv("/KIMEL/tigrlab/scratch/jshah/brain_age_analysis/data/processed/developmental_brainage_output/output_file_2024-10-22-22-50-10_Adjusted_MR_predicted_age_male.csv", show_col_types = FALSE)

Adjusted_BrainAGE_male <- read_csv("/KIMEL/tigrlab/scratch/jshah/brain_age_analysis/data/processed/developmental_brainage_output/output_file_2024-10-22-22-50-10_Adjusted_BrainAGE_male.csv", show_col_types = FALSE)

#Female data
MR_predicted_age_female <- read_csv("/KIMEL/tigrlab/scratch/jshah/brain_age_analysis/data/processed/developmental_brainage_output/output_file_2024-10-22-22-45-31_MR_predicted_age_female.csv", show_col_types = FALSE)

BrainAGE_female <- read_csv("/KIMEL/tigrlab/scratch/jshah/brain_age_analysis/data/processed/developmental_brainage_output/output_file_2024-10-22-22-45-31_BrainAGE_female.csv", show_col_types = FALSE)

Adjusted_MR_predicted_age_female <- read_csv("/KIMEL/tigrlab/scratch/jshah/brain_age_analysis/data/processed/developmental_brainage_output/output_file_2024-10-22-22-45-31_Adjusted_MR_predicted_age_female.csv", show_col_types = FALSE)

Adjusted_BrainAGE_female <- read_csv("/KIMEL/tigrlab/scratch/jshah/brain_age_analysis/data/processed/developmental_brainage_output/output_file_2024-10-22-22-45-31_Adjusted_BrainAGE_female.csv", show_col_types = FALSE)

#Load euler's number

euler_data <- read_tsv("/KIMEL/tigrlab/scratch/jshah/brain_age_analysis/data/raw/00_group2_stats_tables/euler.tsv")

#Load male template because it has real age

male_developmental_template <- read_csv("/KIMEL/tigrlab/scratch/jshah/brain_age_analysis/data/processed/male_subID_developmental_template.csv", show_col_types = FALSE)

#Load female template because it has real age and subject ID

female_developmental_template <- read_csv("/KIMEL/tigrlab/scratch/jshah/brain_age_analysis/data/processed/female_subID_developmental_template.csv", show_col_types = FALSE)

#Load TAY demographic data to get PSS

tay_demo <- read_csv("/KIMEL/tigrlab/scratch/jshah/brain_age_analysis/data/raw/tay_mri_demo_summer2024.csv", show_col_types = FALSE)

```
Function to clean demographic subject_IDs

```{r}
clean_subject_ids_demo <- function(df, subject_id) {
  df %>%
    mutate(clean_subject_id = str_extract(!!sym(subject_id), "[0-9]{8}$"))
}
```

Use the function to clean the demographic subject ids

```{r}
TAY_demographics <- clean_subject_ids_demo(tay_demo, "subject_id")

female_developmental_template <- female_developmental_template %>%
  rename(clean_subject_id = subject_id)

male_developmental_template <- male_developmental_template %>%
  rename(clean_subject_id = subject_id)

```

clean subject IDs for the euler_data file

```{r}
clean_subject_ids_tsv <- function(df) {
  colnames(df)[1] <- "clean_subject_id"  # Rename the first column to 'clean_subject_id'
  df %>%
    mutate(clean_subject_id = str_extract(clean_subject_id, "[0-9]{8}$"))  # Extract the last 8 digits of subject IDs
}

euler_data_clean <- clean_subject_ids_tsv(euler_data)

print(euler_data_clean)
```


filtered for participant rows that have prodromal psychosis because not all participants have prodromal psychosis listed at T1

```{r}
# Group by subject ID and prioritize rows where symptoms are filled out
# If symptoms are NA, the function will pick the first occurrence that has them filled
TAY_unique_symptoms <- TAY_demographics %>%
  group_by(clean_subject_id) %>%
  arrange(desc(!is.na(prodromal_psychosis))) %>%  # Arrange to prioritize rows where symptoms are NOT NA
  slice(1) %>%  # Keep the first row per subject (the one with symptoms, if available)
  ungroup()

# The result will be a dataframe with unique subject IDs and prioritized symptom data
print(TAY_unique_symptoms)
```

Create a new dataframe with subject IDs from the female template, and age. Then import the PSS/non-PSS from the TAY demographics using the matching subject IDs. 

```{r}
# Step 1: Extract subject ID and age from female_developmental_template
female_data <- female_developmental_template %>%
  select(clean_subject_id, AGE)  # Assuming 'clean_subject_id' and 'age' are the column names

# Step 2: Use left_join to get sex and PSS from TAY_demographics
# Assuming 'TAY_demographics' has columns 'clean_subject_id', 'assigned_sex_at_birth', and 'prodromal_psychosis'
final_analysis_female_data <- female_data %>%
  left_join(TAY_unique_symptoms %>% 
              select(clean_subject_id, assigned_sex_at_birth, prodromal_psychosis),  # Select relevant columns
            by = "clean_subject_id")  # Merge by subject ID

print(final_analysis_female_data)
```

Do the same for the male participants 

```{r}
# Step 1: Extract subject ID and age from female_developmental_template
male_data <- male_developmental_template %>%
  select(clean_subject_id, AGE)  # Assuming 'clean_subject_id' and 'age' are the column names

# Step 2: Use left_join to get sex and PSS from TAY_demographics
# Assuming 'TAY_demographics' has columns 'clean_subject_id', 'assigned_sex_at_birth', and 'prodromal_psychosis'
final_analysis_male_data <- male_data %>%
  left_join(TAY_unique_symptoms %>% 
              select(clean_subject_id, assigned_sex_at_birth, prodromal_psychosis),  # Select relevant columns
            by = "clean_subject_id")  # Merge by subject ID

print(final_analysis_male_data)
```


Add the brain age calculator output. This includes female_age, female_error, corage_female, female_adjerror

```{r}
final_analysis_female_data<- final_analysis_female_data %>%
  mutate(female_age = MR_predicted_age_female$female_age, 
         female_error = BrainAGE_female$female_error,
         corage_female = Adjusted_MR_predicted_age_female$corage_female,
         female_adjerror = Adjusted_BrainAGE_female$female_adjerror)

# Rename the columns
final_analysis_female_data <- final_analysis_female_data %>%
  rename(
    predicted_age = female_age, 
    BrainAGE = female_error,
    Adjusted_predicted_age = corage_female,
    Adjusted_BrainAGE = female_adjerror
  )

# Print the updated dataframe to check the new column names
print(final_analysis_female_data)
```

Add brain age calculations for male data 

```{r}
final_analysis_male_data<- final_analysis_male_data %>%
  mutate(male_age = MR_predicted_age_male$male_age, 
         male_error = BrainAGE_male$male_error,
         corage_male = Adjusted_MR_predicted_age_male$corage_male,
         male_adjerror = Adjusted_BrainAGE_male$male_adjerror)

# Rename the columns
final_analysis_male_data <- final_analysis_male_data %>%
  rename(
    predicted_age = male_age, 
    BrainAGE = male_error,
    Adjusted_predicted_age = corage_male,
    Adjusted_BrainAGE = male_adjerror
  )

print(final_analysis_male_data)
```

combine the female and male data to create one dataframe to work with

```{r}
combined_analysis_data <- rbind(final_analysis_female_data, final_analysis_male_data)

print(combined_analysis_data)
```

Add euler's number to the combined_analysis_data and join based on subject ID

```{r}
combined_analysis_data <- combined_analysis_data %>%
  left_join(euler_data_clean %>% 
              select(clean_subject_id, mean_euler_bh),  # Select relevant columns
            by = "clean_subject_id")  # Merge by subject ID

print(combined_analysis_data)
```


Run a correlational analysis between AGE (real age) and corage_female 

```{r}

plot(combined_analysis_data$AGE, combined_analysis_data$Adjusted_predicted_age, main = "Association Between Real Age and Predicted Age",
     xlab = "Real Age", ylab = "Predicted Age",
     pch = 19, frame = FALSE)
cor.test(combined_analysis_data$AGE, combined_analysis_data$Adjusted_predicted_age, method = 'pearson')
  
```


Run linear regression between brain age and psychosis symptoms

```{r}

plot(combined_analysis_data$prodromal_psychosis, combined_analysis_data$Adjusted_predicted_age,
     main = "Association Between Psychosis Spectrum Symptoms and Predicted BrainAGE",
     xlab = "Psychosis Spectrum Symptoms", ylab = "Predicted Age",
     pch = 19, frame = FALSE, col = "blue")

# Run a linear regression analysis
lm_model <- lm(Adjusted_predicted_age ~ prodromal_psychosis, data = combined_analysis_data)

# Add the regression line to the plot
abline(lm_model, col = "red", lwd = 2)

# Print the linear regression model summary
summary(lm_model)

```

Run regression analysis for BrainAGE and PSS

```{r}
plot(combined_analysis_data$prodromal_psychosis, combined_analysis_data$Adjusted_BrainAGE,
     main = "Association Between Psychosis Spectrum Symptoms and Predicted BrainAGE",
     xlab = "Psychosis Spectrum Symptoms", ylab = "Adjusted BrainAGE",
     pch = 19, frame = FALSE, col = "blue")

# Run a linear regression analysis
lm_model <- lm(prodromal_psychosis ~ Adjusted_BrainAGE, data = combined_analysis_data)

# Add the regression line to the plot
abline(lm_model, col = "red", lwd = 2)

# Print the linear regression model summary
summary(lm_model)
```

Run multivariable regression with PSS as the independent variable and predicted brainage as the dependent variable

```{r}
lm_model <- lm(Adjusted_predicted_age ~ prodromal_psychosis + assigned_sex_at_birth + mean_euler_bh, data = combined_analysis_data)

# View the summary of the regression model
summary(lm_model)

# Plot the relationship between predicted age and PSS
plot(combined_analysis_data$prodromal_psychosis, combined_analysis_data$Adjusted_predicted_age,
     main = "Association Between Predicted Age and PSS",
     xlab = "Psychosis Spectrum Symptoms", ylab = "Predicted Age",
     pch = 19, frame = FALSE,)


```
Multivariable regression for BrainAGE and PSS

```{r}
lm_model <- lm(Adjusted_BrainAGE ~ prodromal_psychosis + assigned_sex_at_birth + mean_euler_bh, data = combined_analysis_data)

# View the summary of the regression model
summary(lm_model)

# Plot the relationship between predicted age and PSS
plot(combined_analysis_data$prodromal_psychosis, combined_analysis_data$Adjusted_BrainAGE, 
     main = "Association Between BrainAGE and PSS",
     xlab ="Psychosis Spectrum Symptoms", ylab = "Adjusted BrainAGE" ,
     pch = 19, frame = FALSE,)

```
Plot the variables of interest just to get an idea of what the data looks like

```{r}
# Plot the relationship between predicted age and Mean Euler
plot(combined_analysis_data$Adjusted_BrainAGE, combined_analysis_data$mean_euler_bh,
     main = "Association Between BrainAGE and Mean Euler",
     xlab = "Adjusted BrainAGE", ylab = "Mean Euler",
     pch = 19, frame = FALSE,)

# Plot the relationship between predicted PSS and Sex
plot(combined_analysis_data$prodromal_psychosis, combined_analysis_data$Adjusted_BrainAGE, 
     main = "Association Between PSS and Sex",
     xlab = "PSS", ylab = "Sex",
     pch = 19, frame = FALSE,)

```


```{r}
