---
title: "developmental_brainage_analysis"
output: html_document
---

load libraries

```{r}
library(ggplot2)
library(readr)
library(readxl)
library(rstatix)
library(broom)
library(stringr)
library(dplyr)
```

load files

```{r}

#Male data
MR_predicted_age_male <- read_csv("/KIMEL/tigrlab/scratch/jshah/brain_age_analysis/data/processed/full_developmental_output/output_file_2024-10-28-00-45-29_MR_predicted_age_male.csv", show_col_types = FALSE)

BrainAGE_male <- read_csv("/KIMEL/tigrlab/scratch/jshah/brain_age_analysis/data/processed/full_developmental_output/output_file_2024-10-28-00-45-29_BrainAGE_male.csv", show_col_types = FALSE)

Adjusted_MR_predicted_age_male <- read_csv("/KIMEL/tigrlab/scratch/jshah/brain_age_analysis/data/processed/full_developmental_output/output_file_2024-10-28-00-45-29_Adjusted_MR_predicted_age_male.csv", show_col_types = FALSE)

Adjusted_BrainAGE_male <- read_csv("/KIMEL/tigrlab/scratch/jshah/brain_age_analysis/data/processed/full_developmental_output/output_file_2024-10-28-00-45-29_Adjusted_BrainAGE_male.csv", show_col_types = FALSE)

#Female data
MR_predicted_age_female <- read_csv("/KIMEL/tigrlab/scratch/jshah/brain_age_analysis/data/processed/full_developmental_output/output_file_2024-10-28-00-40-48_MR_predicted_age_female.csv", show_col_types = FALSE)

BrainAGE_female <- read_csv("/KIMEL/tigrlab/scratch/jshah/brain_age_analysis/data/processed/full_developmental_output/output_file_2024-10-28-00-40-48_BrainAGE_female.csv", show_col_types = FALSE)

Adjusted_MR_predicted_age_female <- read_csv("/KIMEL/tigrlab/scratch/jshah/brain_age_analysis/data/processed/full_developmental_output/output_file_2024-10-28-00-40-48_Adjusted_MR_predicted_age_female.csv", show_col_types = FALSE)

Adjusted_BrainAGE_female <- read_csv("/KIMEL/tigrlab/scratch/jshah/brain_age_analysis/data/processed/full_developmental_output/output_file_2024-10-28-00-40-48_Adjusted_BrainAGE_female.csv", show_col_types = FALSE)

#Load euler's number

euler_data <- read_tsv("/KIMEL/tigrlab/scratch/jshah/brain_age_analysis/data/raw/00_group2_stats_tables/euler.tsv")

#Load male template because it has real age

male_developmental_template <- read_csv("/KIMEL/tigrlab/scratch/jshah/brain_age_analysis/data/processed/male_subID_developmental_template.csv", show_col_types = FALSE)

#Load female template because it has real age and subject ID

female_developmental_template <- read_csv("/KIMEL/tigrlab/scratch/jshah/brain_age_analysis/data/processed/female_subID_developmental_template.csv", show_col_types = FALSE)

#Load TAY demographic data to get PSS

tay_demo <- read_csv("/KIMEL/tigrlab/scratch/jshah/brain_age_analysis/data/raw/tay_mri_demo_summer2024.csv", show_col_types = FALSE)

#Load data cut file 
datacut_ids <- read_csv("/KIMEL/tigrlab/scratch/jshah/brain_age_analysis/data/raw/datacut_ids.csv")

```
Function to clean demographic subject_IDs

```{r}
clean_subject_ids_demo <- function(df, subject_id) {
  df %>%
    mutate(clean_subject_id = str_extract(!!sym(subject_id), "[0-9]{8}$"))
}
```

Use the function to clean the demographic subject ids

```{r}
TAY_demographics <- clean_subject_ids_demo(tay_demo, "subject_id")

datacut_ids_clean <- clean_subject_ids_demo(datacut_ids, "subject_id")

female_developmental_template <- female_developmental_template %>%
  rename(clean_subject_id = subject_id)

female_developmental_template <- female_developmental_template %>%
  mutate(sex = "Female")

male_developmental_template <- male_developmental_template %>%
  rename(clean_subject_id = subject_id)

male_developmental_template <- male_developmental_template %>%
  mutate(sex = "Male")

```

clean subject IDs for the euler_data file

```{r}
clean_subject_ids_tsv <- function(df) {
  colnames(df)[1] <- "clean_subject_id"  # Rename the first column to 'clean_subject_id'
  df %>%
    mutate(clean_subject_id = str_extract(clean_subject_id, "[0-9]{8}$"))  # Extract the last 8 digits of subject IDs
}

euler_data_clean <- clean_subject_ids_tsv(euler_data)

print(euler_data_clean)
```


filtered for participant rows that have prodromal psychosis because not all participants have prodromal psychosis listed at T1

```{r}
# Group by subject ID and prioritize rows where symptoms are filled out
# If symptoms are NA, the function will pick the first occurrence that has them filled
TAY_unique_symptoms <- TAY_demographics %>%
  group_by(clean_subject_id) %>%
  arrange(desc(!is.na(prodromal_psychosis))) %>%  # Arrange to prioritize rows where symptoms are NOT NA
  slice(1) %>%  # Keep the first row per subject (the one with symptoms, if available)
  ungroup()

# The result will be a dataframe with unique subject IDs and prioritized symptom data
print(TAY_unique_symptoms)
```

Create a new dataframe with subject IDs from the female template, and age. Then import the PSS/non-PSS from the TAY demographics using the matching subject IDs. 

```{r}
# Step 1: Extract subject ID and age from female_developmental_template
female_data <- female_developmental_template %>%
  select(clean_subject_id, AGE, sex)  # Assuming 'clean_subject_id' and 'age' are the column names

# Step 2: Use left_join to get sex and PSS from TAY_demographics
# Assuming 'TAY_demographics' has columns 'clean_subject_id', 'assigned_sex_at_birth', and 'prodromal_psychosis'
final_analysis_female_data <- female_data %>%
  left_join(TAY_unique_symptoms %>% 
              select(clean_subject_id, prodromal_psychosis),  # Select relevant columns
            by = "clean_subject_id")  # Merge by subject ID

print(final_analysis_female_data)
```

Do the same for the male participants 

```{r}
# Step 1: Extract subject ID and age from female_developmental_template
male_data <- male_developmental_template %>%
  select(clean_subject_id, AGE, sex)  # Assuming 'clean_subject_id' and 'age' are the column names

# Step 2: Use left_join to get sex and PSS from TAY_demographics
# Assuming 'TAY_demographics' has columns 'clean_subject_id', 'assigned_sex_at_birth', and 'prodromal_psychosis'
final_analysis_male_data <- male_data %>%
  left_join(TAY_unique_symptoms %>% 
              select(clean_subject_id, prodromal_psychosis),  # Select relevant columns
            by = "clean_subject_id")  # Merge by subject ID

print(final_analysis_male_data)
```


Add the brain age calculator output. This includes female_age, female_error, corage_female, female_adjerror

```{r}
final_analysis_female_data<- final_analysis_female_data %>%
  mutate(female_age = MR_predicted_age_female$female_age, 
         female_error = BrainAGE_female$female_error,
         corage_female = Adjusted_MR_predicted_age_female$corage_female,
         female_adjerror = Adjusted_BrainAGE_female$female_adjerror)

# Rename the columns
final_analysis_female_data <- final_analysis_female_data %>%
  rename(
    predicted_age = female_age, 
    BrainAGE = female_error,
    Adjusted_predicted_age = corage_female,
    Adjusted_BrainAGE = female_adjerror
  )

# Print the updated dataframe to check the new column names
print(final_analysis_female_data)
```

Add brain age calculations for male data 

```{r}
final_analysis_male_data<- final_analysis_male_data %>%
  mutate(male_age = MR_predicted_age_male$male_age, 
         male_error = BrainAGE_male$male_error,
         corage_male = Adjusted_MR_predicted_age_male$corage_male,
         male_adjerror = Adjusted_BrainAGE_male$male_adjerror)

# Rename the columns
final_analysis_male_data <- final_analysis_male_data %>%
  rename(
    predicted_age = male_age, 
    BrainAGE = male_error,
    Adjusted_predicted_age = corage_male,
    Adjusted_BrainAGE = male_adjerror
  )

print(final_analysis_male_data)
```

combine the female and male data to create one dataframe to work with

```{r}
combined_analysis_data <- rbind(final_analysis_female_data, final_analysis_male_data)

print(combined_analysis_data)
```

Add euler's number to the combined_analysis_data and join based on subject ID

```{r}
combined_analysis_data <- combined_analysis_data %>%
  left_join(euler_data_clean %>% 
              select(clean_subject_id, mean_euler_bh),  # Select relevant columns
            by = "clean_subject_id")  # Merge by subject ID

print(combined_analysis_data)
```

Filter for people who are in the TAY data cut 

```{r}
# Assuming 'sub_ID' is the column to match in both data frames and that `t01` is a logical or character column.
combined_analysis_data <- combined_analysis_data %>%
  # Join with datacut_ids_clean to bring in the 't01' column
  left_join(datacut_ids_clean %>% select(clean_subject_id, t01), by = "clean_subject_id")

# View the result
print(combined_analysis_data)

datacut_filtered_data <- combined_analysis_data %>%
  filter(t01 == TRUE)

# View the result
print(datacut_filtered_data)

```
Remove high euler number (cut off above 200)

```{r}
euler_filtered_data <- datacut_filtered_data %>%
  filter(mean_euler_bh >= -200)

print(euler_filtered_data)

```


View all NAs

```{r}
rows_with_na_develop <- euler_filtered_data %>%
  filter(if_any(everything(), is.na))
sum(is.na(euler_filtered_data$prodromal_psychosis))

print(rows_with_na_develop)

cleaned_filtered_data <- na.omit(euler_filtered_data)

print(cleaned_filtered_data)

write.csv(cleaned_filtered_data, "/KIMEL/tigrlab/scratch/jshah/brain_age_analysis/data/processed/complete_developmental_data.csv", row.names = FALSE)
```


Run summary statistics

```{r}

summary(cleaned_filtered_data)
table(cleaned_filtered_data$sex)

```

Run a correlational analysis between AGE (real age) and corage_female 

```{r}

plot(cleaned_filtered_data$AGE, cleaned_filtered_data$Adjusted_predicted_age, main = "Association Between Real Age and Adjusted Predicted Age",
     xlab = "Real Age", ylab = "Adjusted Predicted Age",
     pch = 19, frame = FALSE, col = "blue")
cor.test(cleaned_filtered_data$AGE, cleaned_filtered_data$Adjusted_predicted_age, method = 'pearson')

plot(cleaned_filtered_data$AGE, cleaned_filtered_data$predicted_age, main = "Association Between Real Age and Predicted Age",
     xlab = "Real Age", ylab = "Predicted Age",
     pch = 19, frame = FALSE)
cor.test(cleaned_filtered_data$AGE, cleaned_filtered_data$predicted_age, method = 'pearson')

```


Run linear regression between predicted age and psychosis symptoms

```{r}

plot(cleaned_filtered_data$prodromal_psychosis, cleaned_filtered_data$Adjusted_predicted_age,
     main = "Association Between Psychosis Spectrum Symptoms and Predicted BrainAGE",
     xlab = "Psychosis Spectrum Symptoms", ylab = "Predicted Age",
     pch = 19, frame = FALSE, col = "blue")

# Run a linear regression analysis
lm_model <- lm(Adjusted_predicted_age ~ prodromal_psychosis, data = cleaned_filtered_data)

# Add the regression line to the plot
abline(lm_model, col = "red", lwd = 2)

# Print the linear regression model summary
summary(lm_model)

```

Run regression analysis for BrainAGE and PSS

```{r}
plot(cleaned_filtered_data$prodromal_psychosis, cleaned_filtered_data$Adjusted_BrainAGE,
     main = "Association Between Psychosis Spectrum Symptoms and Predicted BrainAGE",
     xlab = "Psychosis Spectrum Symptoms", ylab = "Adjusted BrainAGE",
     pch = 19, frame = FALSE, col = "blue")

# Run a linear regression analysis
lm_model <- lm(prodromal_psychosis ~ Adjusted_BrainAGE, data = cleaned_filtered_data)

# Add the regression line to the plot
abline(lm_model, col = "red", lwd = 2)

# Print the linear regression model summary
summary(lm_model)
```

Run multivariable regression with PSS as the independent variable and predicted brainage as the dependent variable

```{r}
run_lm_model <- function(dependent_var, independent_vars, data) {
  # Create a formula dynamically
  formula <- as.formula(paste(dependent_var, "~", paste(independent_vars, collapse = " + ")))

  # Run the linear model with the formula and the data explicitly specified
  lm_model <- lm(formula, data = data)

  # Print summary of the model
  return(summary(lm_model))
}

dependent_vars <- c("Adjusted_predicted_age", "Adjusted_BrainAGE", "predicted_age", "BrainAGE")

for (dep_var in dependent_vars) {
  cat("Running linear model for:", dep_var, "\n")
  lm_summary <- run_lm_model(dep_var, c("prodromal_psychosis", "sex", "mean_euler_bh"), cleaned_filtered_data)
  print(lm_summary)  # Print the summary of the model
}

# Plot the relationship between predicted age and PSS
plot(cleaned_filtered_data$prodromal_psychosis, cleaned_filtered_data$Adjusted_BrainAGE, 
     main = "Association Between BrainAGE and PSS",
     xlab ="Psychosis Spectrum Symptoms", ylab = "Adjusted BrainAGE" ,
     pch = 19, frame = FALSE,)

```
Plot the variables of interest just to get an idea of what the data looks like

```{r}
# Plot the relationship between predicted age and Mean Euler (nothing removed)
plot(datacut_filtered_data$Adjusted_BrainAGE, datacut_filtered_data$mean_euler_bh,
     main = "Association Between BrainAGE and Mean Euler",
     xlab = "BrainAGE", ylab = "Mean Euler",
     pch = 19, frame = FALSE,)

# Plot the relationship between predicted age and Mean Euler with high euler removed
plot(cleaned_filtered_data$Adjusted_BrainAGE, cleaned_filtered_data$mean_euler_bh,
     main = "Association Between BrainAGE and Mean Euler",
     xlab = "Adjusted BrainAGE", ylab = "Mean Euler",
     pch = 19, frame = FALSE, col = "blue")

# Plot the relationship between predicted PSS and Sex
plot(cleaned_filtered_data$prodromal_psychosis, cleaned_filtered_data$Adjusted_BrainAGE, 
     main = "Association Between PSS and Adjusted BrainAGE",
     xlab = "PSS", ylab = "Adjusted BrainAGE",
     pch = 19, frame = FALSE,)

```
Calculate effect sizes 

```{r}


effect_size <- cleaned_filtered_data %>% 
  cohens_d(Adjusted_BrainAGE ~ prodromal_psychosis)

effect_size_pred_age <- cleaned_filtered_data %>% 
  cohens_d(predicted_age ~ prodromal_psychosis)

# View the result
print(effect_size)
print(effect_size_pred_age)

```

Plot BrainAGE and PSS
```{r}
# Create a box plot for brain age by PSS status
ggplot(cleaned_filtered_data, aes(x = factor(prodromal_psychosis, labels = c("PSS = FALSE", "PSS = TRUE")), y = Adjusted_BrainAGE)) +
  geom_boxplot(fill = c("lightblue", "salmon")) +  # Add color to the boxes
  labs(title = "Adjusted BrainAGE According to PSS",
       x = "PSS",
       y = "Adjusted BrainAGE") +
  theme_minimal()

# Create a box plot for predicted age by PSS status
ggplot(cleaned_filtered_data, aes(x = factor(prodromal_psychosis, labels = c("PSS = FALSE", "PSS = TRUE")), y = Adjusted_predicted_age)) +
  geom_boxplot(fill = c("lightblue", "salmon")) +  # Add color to the boxes
  labs(title = "Adjusted Predicted Age According to PSS",
       x = "PSS",
       y = "Adjusted Predicted Age") +
  theme_minimal()

```


```{r}
