---
title: "Test_developmental_brain_age"
output: html_notebook
date: 2024-1017
---

Load the necessary libraries

```{r}
library(readxl)
library(openxlsx)
library(stringr)
library(dplyr)
library(readr)
```

Load Files

```{r}
#Load excel file
template_brainage <- read_excel("/KIMEL/tigrlab/scratch/jshah/TAY/freesurfer/output/template_brainage.xlsx")

#Load parcellation files
lh_thick <- read_tsv("/KIMEL/tigrlab/scratch/jshah/TAY/freesurfer/output/00_group2_stats_tables/lh.Schaefer2018_400Parcels.thickness.tsv")
lh_surf <- read_tsv("/KIMEL/tigrlab/scratch/jshah/TAY/freesurfer/output/00_group2_stats_tables/lh.Schaefer2018_400Parcels.surfacearea.tsv")

rh_thick <- read_tsv("/KIMEL/tigrlab/scratch/jshah/TAY/freesurfer/output/00_group2_stats_tables/rh.Schaefer2018_400Parcels.thickness.tsv")
rh_surf <- read_tsv("/KIMEL/tigrlab/scratch/jshah/TAY/freesurfer/output/00_group2_stats_tables/rh.Schaefer2018_400Parcels.surfacearea.tsv")

TAY_demographics <- read_csv("/KIMEL/tigrlab/scratch/jshah/tay_mri_demo_summer2024.csv")
```

Step 2: Function to clean column names by removing underscores and removing first n characters
```{r}
#save the original column names
original_template_column_names <- colnames(template_brainage)

#function to clean columns
clean_columns <- function(df, remove_char_count) {
  if (is.data.frame(df) && ncol(df) > 0) {
    colnames(df) <- str_replace_all(colnames(df), "_", "")  # Remove underscores
 
    # Check length of each column name and conditionally remove characters
    colnames(df) <- ifelse(nchar(colnames(df)) > remove_char_count,
                           str_sub(colnames(df), remove_char_count + 1), 
                           colnames(df))  # Remove first n characters if long enough
  }
  return(df)
}
   
```

Clean the dataframes using the function

```{r}
lh_thickness_df <- clean_columns(lh_thick, 11)  # Clean lh_thick first
lh_area_df <- clean_columns(lh_surf, 11)  # Clean lh_surf
rh_thickness_df <- clean_columns(rh_thick, 11)  # Clean rh_thick first
rh_area_df <- clean_columns(rh_surf, 11)  # Clean rh_surf

template_df <- clean_columns(template_brainage, 10)
```

Rename thickness_df from thickness -> to thick

```{r}
colnames(lh_thickness_df) <- str_replace_all(colnames(lh_thickness_df), "thickness", "thick") 
colnames(rh_thickness_df) <- str_replace_all(colnames(rh_thickness_df), "thickness", "thick") 

```

Step 3: Function to clean and standardize subject IDs for lh_thick and lh_surf

```{r}
clean_subject_ids_tsv <- function(df) {
  colnames(df)[1] <- "clean_subject_id"  # Rename the first column to 'clean_subject_id'
  df %>%
    mutate(clean_subject_id = str_extract(clean_subject_id, "[0-9]{8}$"))  # Extract the last 8 digits of subject IDs
}
```

Apply the subject_ID cleaning function to thickness and surface area files

```{r}
lh_thick_clean <- clean_subject_ids_tsv(lh_thickness_df)
lh_surf_clean <- clean_subject_ids_tsv(lh_area_df)
rh_thick_clean <- clean_subject_ids_tsv(rh_thickness_df)
rh_surf_clean <- clean_subject_ids_tsv(rh_area_df)
```

Step 3b: Function to clean demo subject_IDs

```{r}
clean_subject_ids_demo <- function(df, subject_id) {
  df %>%
    mutate(clean_subject_id = str_extract(!!sym(subject_id), "[0-9]{8}$"))
}
```

Clean subject IDs (after renaming columns)

```{r}
TAY_demographics <- clean_subject_ids_demo(TAY_demographics, "subject_id")
```

Step 3c: Rename subject_id to clean_subject_id in the 

```{r}
template_df <- template_df %>%
  rename(clean_subject_id = subjectid)
```

Step 4: Filter for female subjects only from demographics dataframe

```{r}
female_data <- TAY_demographics %>%
  filter(assigned_sex_at_birth == "Female") %>%
  select(clean_subject_id, age_scan)  # Keep subject ID and age for merging
```

Step 5: Filter the thickness and area data for female subject IDs -> create function to filter subject_IDs? 

```{r}
female_lh_thickness_data <- lh_thick_clean %>%
  filter(clean_subject_id %in% female_data$clean_subject_id)

female_lh_area_data <- lh_surf_clean %>%
  filter(clean_subject_id %in% female_data$clean_subject_id)

female_rh_thickness_data <- rh_thick_clean %>%
  filter(clean_subject_id %in% female_data$clean_subject_id)

female_rh_area_data <- rh_surf_clean %>%
  filter(clean_subject_id %in% female_data$clean_subject_id)
```

make the characters the same
```{r}
template_df$clean_subject_id <-as.character(template_df$clean_subject_id)
```

Ensure that all columns remain intact while adding the female data

```{r}
# First, check if template_df has rows
if (nrow(template_df) == 0) {
  # Create an empty data frame with the same structure as template_df but add the subject IDs and age
  # Keep all the original columns, but add the new data
  template_df <- template_df[0, ]  # Retain column structure but no rows
  
  # Now create a new data frame that only updates clean_subject_id and AGE
  new_data <- data.frame(clean_subject_id = female_data$clean_subject_id,
                         AGE = female_data$age_scan,
                         stringsAsFactors = FALSE)
  
  # Bind the new data to the empty template, ensuring columns remain intact
  template_df <- bind_rows(template_df, new_data)
} else {
  # If template_df already has rows, just update clean_subject_id and AGE columns
  template_df$clean_subject_id <- female_data$clean_subject_id
  template_df$AGE <- female_data$age_scan
}

```


Check if subject IDs from both dataframes match

```{r}
common_ids <- intersect(template_df$clean_subject_id, female_lh_thickness_data$clean_subject_id)
print(paste("Number of matching subject IDs:", length(common_ids)))
```

Make all of the logical variables into characters
```{r}

# Convert all columns in the template_df to characters
template_df <- template_df %>%
  mutate(across(everything(), as.character))

female_data <- female_data %>%
  mutate(across(everything(), as.character))

female_lh_thickness_data <- female_lh_thickness_data %>%
  mutate(across(everything(), as.character))

female_lh_area_data <- female_lh_area_data %>%
  mutate(across(everything(), as.character))

female_rh_thickness_data <- female_rh_thickness_data %>%
  mutate(across(everything(), as.character))

female_rh_area_data <- female_rh_area_data %>%
  mutate(across(everything(), as.character))
```

Step 6: Populate the template with female subject data
Ensure that all the relevant data is properly merged

```{r}
final_template <- left_join(template_df[, c(1, 2)], female_lh_thickness_data, by = "clean_subject_id") %>%  
  left_join(female_lh_area_data, by = "clean_subject_id") %>%
  left_join(female_rh_area_data, by = "clean_subject_id") %>%
  left_join(female_rh_thickness_data, by = "clean_subject_id")

# Select only the columns from template_df
final_template <- final_template %>%
  select(names(template_df))  # Keeps only the columns in template_df  

#remove columns with NA
final_template <- final_template %>%
  filter(!is.na(final_template[[3]]))

#Revert back to original column names found in the template
colnames(final_template)[1:length(original_template_column_names)] <- original_template_column_names
```

Try to change all the types to number
```{r}
# Extract the first column
first_column <- final_template[, 1]

# Convert all other rows and columns to numeric, starting from row 2 and column 2
numeric_data <- final_template[, -1] %>%
  mutate_all(~suppressWarnings(as.numeric(as.character(.))))

```

Step 7: Save the final populated template into a new Excel file

```{r}
output_file <- ("/KIMEL/tigrlab/scratch/jshah/TAY/freesurfer/output/populated_brain_age_template.xlsx")
write.xlsx(numeric_data, output_file)

print("Template has been populated and saved as 'populated_brain_age_template.xlsx'.")
```

