---
title: "Network_brain_age"
output: html_notebook
date: 2024-10-18
---
This is a script to create the Developmental BrainAge template (https://centilebrain.org/#/brainAGE) using freesurfer outputs.


Load the necessary libraries

```{r}
library(readxl)
library(openxlsx)
library(stringr)
library(dplyr)
library(readr)
```

Load Files.
Change sex to "Male" or "Female" depending on analysis.

```{r}

#Load excel file
template_brainage <- read_excel("../data/raw/Network_BrainAGE_template.xlsx")

sex <- "Female"

#Load parcellation files
lh_thick <- read_tsv("../data/raw/00_group2_stats_tables/lh.Schaefer2018_1000Parcels.thickness.tsv")
lh_surf <- read_tsv("../data/raw/00_group2_stats_tables/lh.Schaefer2018_1000Parcels.surfacearea.tsv")

rh_thick <- read_tsv("../data/raw/00_group2_stats_tables/rh.Schaefer2018_1000Parcels.thickness.tsv")
rh_surf <- read_tsv("../data/raw/00_group2_stats_tables/rh.Schaefer2018_1000Parcels.surfacearea.tsv")

TAY_demographics <- read_csv("../data/raw/tay_mri_demo_summer2024.csv")
```

Function to clean and standardize subject IDs for the thickness and surface area data

```{r}
clean_subject_ids_tsv <- function(df) {
  colnames(df)[1] <- "clean_subject_id"  # Rename the first column to 'clean_subject_id'
  df %>%
    mutate(clean_subject_id = str_extract(clean_subject_id, "[0-9]{8}$"))  # Extract the last 8 digits of subject IDs
}
```

Apply the subject_ID cleaning function to thickness and surface area files

```{r}
lh_thick_clean <- clean_subject_ids_tsv(lh_thickness_df)
lh_surf_clean <- clean_subject_ids_tsv(lh_area_df)
rh_thick_clean <- clean_subject_ids_tsv(rh_thickness_df)
rh_surf_clean <- clean_subject_ids_tsv(rh_area_df)
```

Function to clean demographic subject_IDs

```{r}
clean_subject_ids_demo <- function(df, subject_id) {
  df %>%
    mutate(clean_subject_id = str_extract(!!sym(subject_id), "[0-9]{8}$"))
}
```

Use the function to clean the demographic subject ids

```{r}
TAY_demographics <- clean_subject_ids_demo(TAY_demographics, "subject_id")
```

Rename subject_id to clean_subject_id in the demographic file (to standardize the column name)

```{r}
template_df <- template_brainage %>%
  rename(clean_subject_id = SubjectID)
```

Filter for sex specific subjects only from demographics dataframe and also take out age

```{r}
gendered_data <- TAY_demographics %>%
  filter(assigned_sex_at_birth == sex) %>%
  select(clean_subject_id, age_scan, assigned_sex_at_birth)  # Keep subject ID, age, and sex for merging
```

Filter the thickness and area data for female subject IDs -> create function to filter subject_IDs? 

```{r}
gendered_lh_thickness_data <- lh_thick_clean %>%
  filter(clean_subject_id %in% gendered_data$clean_subject_id)

gendered_lh_area_data <- lh_surf_clean %>%
  filter(clean_subject_id %in% gendered_data$clean_subject_id)

gendered_rh_thickness_data <- rh_thick_clean %>%
  filter(clean_subject_id %in% gendered_data$clean_subject_id)

gendered_rh_area_data <- rh_surf_clean %>%
  filter(clean_subject_id %in% gendered_data$clean_subject_id)
```

Ensure that all columns remain intact while adding the female data

```{r}

# Make sure the subject_ids, sex, and age are character data types
template_df$clean_subject_id <- as.character(template_df$clean_subject_id)
template_df$sex <- as.character(template_df$sex)  # Add sex column conversion
template_df$age <- as.character(template_df$age)  # Add age column conversion

# First, check if template_df has rows
if (nrow(template_df) == 0) {
  # Create an empty data frame with the same structure as template_df but add the subject IDs and age
  # Keep all the original columns, but add the new data
  template_df <- template_df[0, ]  # Retain column structure but no rows
  
  # Now create a new data frame that only updates clean_subject_id and AGE
  new_data <- data.frame(clean_subject_id =gendered_data$clean_subject_id,
                         age = gendered_data$age_scan,
                         sex = gendered_data$assigned_sex_at_birth,
                         stringsAsFactors = FALSE)
  
  # Make them the same data type
  new_data$clean_subject_id <- as.character(new_data$clean_subject_id)
  new_data$sex <- as.character(new_data$sex)
  new_data$age <- as.character(new_data$age)
  
  # Bind the new data to the empty template, ensuring columns remain intact
  template_df <- bind_rows(template_df, new_data)
} else {
  
  # If template_df already has rows, just update clean_subject_id and AGE columns
  template_df$clean_subject_id <- gendered_data$clean_subject_id
  template_df$AGE <- gendered_data$age_scan
}

```


Check if subject IDs from both dataframes match

```{r}
common_ids <- intersect(template_df$clean_subject_id, gendered_lh_thickness_data$clean_subject_id)
print(paste("Number of matching subject IDs:", length(common_ids)))
```

Make all of the logical variables into characters

```{r}

# Convert all columns in the template_df to characters
template_df <- template_df %>%
  mutate(across(everything(), as.character))

gendered_data <- gendered_data %>%
  mutate(across(everything(), as.character))

gendered_lh_thickness_data <- gendered_lh_thickness_data %>%
  mutate(across(everything(), as.character))

gendered_lh_area_data <- gendered_lh_area_data %>%
  mutate(across(everything(), as.character))

gendered_rh_thickness_data <- gendered_rh_thickness_data %>%
  mutate(across(everything(), as.character))

gendered_rh_area_data <- gendered_rh_area_data %>%
  mutate(across(everything(), as.character))
```

Populate the template with female subject data and only keep the column names that are in the template. Remove any rows with NA, the calculator will not work if there are any NA values. Then revert back to the original column names. 
Ensure that all the relevant data is properly merged.

```{r}
final_template <- left_join(template_df[, c(1, 2)], gendered_lh_thickness_data, by = "clean_subject_id") %>%  
  left_join(gendered_lh_area_data, by = "clean_subject_id") %>%
  left_join(gendered_rh_thickness_data, by = "clean_subject_id") %>%
  left_join(gendered_rh_area_data, by = "clean_subject_id")

# Select only the columns from template_df
final_template <- final_template %>%
  select(names(template_df))  # Keeps only the columns in template_df  

#remove columns with NA
final_template <- final_template %>%
  filter(!is.na(final_template[[3]]))

```

Change all the data types to numeric and remove the first column of the subject ids because that was not in the original template. Make sure all the data types are numeric because the calculator will not work if they are different data types. 

```{r}
# Extract the first column
first_column <- final_template[, 1]

# Convert all other rows and columns to numeric, starting from row 2 and column 2
numeric_data <- final_template[, -1] %>%
  mutate_all(~suppressWarnings(as.numeric(as.character(.))))

```

Save the final populated template into a new Excel file

```{r}

output_file <- paste0("../data/processed/", tolower(sex), "_populated_developmental_template.xlsx")
write.xlsx(numeric_data, output_file)

print(paste("Template has been populated and saved as", output_file))
```

