---
title: "network_brainage_merge"
output:
  html_document: default
  pdf_document: default
---
Network BrainAGE Analysis Script: This script will combine brainAGE output files and run statistical analysis on the brainAGE outputs and PSS. 

load libraries

```{r}
library(ggplot2)
library(readr)
library(readxl)
library(rstatix)
library(broom)
library(stringr)
library(dplyr)
```

The network brain age templated was segmented by 150 participants and thus, the output files are also broken up in this way. To simplify the analysis, the files need to be merged into one .csv file. 

Write a function that will merge files with a specific pattern and output a merged file

```{r}
merge_files_with_pattern <- function(directory, common_patterns, output_file_name) {
  for (common_pattern in common_patterns) {
      # Create the file pattern using the provided common part of the file names
    pattern <- paste0(".*", common_pattern, ".*\\.csv")
  
    # List all files in the directory that match the pattern
    files <- list.files(path = directory, pattern = pattern, full.names = TRUE)
  
    # Check if there are matching files
    if (length(files) == 0) {
      message(paste("No files found with the specified pattern:", common_pattern))
      next
    }
    
        # Define the output file path
    output_path <- file.path(directory, paste0("combined_", common_pattern, ".csv"))
    
    # Check if the output file already exists
    if (file.exists(output_path)) {
      message(paste("File already exists:", output_path, "- skipping file creation."))
      next
    }
    
    # Initialize an empty dataframe to store the combined data
    combined_data <- data.frame()
    
    # Loop through each file, read its content, and bind the rows
    for (file in files) {
      # Read the data from the CSV file
      file_data <- read.csv(file)
      
      # Combine the data into the final dataframe
      combined_data <- rbind(combined_data, file_data)
    }
    
    # Save the combined data into a new CSV file
    write.csv(combined_data, output_path, row.names = FALSE)
    
    # Print a success message
    print(paste("All files with pattern", common_pattern, "have been merged and saved as:", output_path))
  }
}

```

Create a list of the file patterns that need to be put into the function and then loop the function through each file pattern to create 8 new merged files with network brainage outputs for both males and females.

```{r}
directory <- "/KIMEL/tigrlab/scratch/jshah/brain_age_analysis/data/processed/Full_Network_output"
common_patterns <- c("Network_Adjusted_BrainAGE", "Network_MR_predicted_age", "Network_BrainAGE", "Network_Adjusted_MR_predicted_age")
output_file_prefix <- "combined_output"
merge_files_with_pattern(directory, common_patterns, output_file_prefox)

```
load necessary files

```{r}

#load network output files
Network_Adjusted_BrainAGE <- read_csv("/KIMEL/tigrlab/scratch/jshah/brain_age_analysis/data/processed/Full_Network_output/combined_Network_Adjusted_BrainAGE.csv", show_col_types = FALSE)

Network_MR_predicted_age <- read_csv("/KIMEL/tigrlab/scratch/jshah/brain_age_analysis/data/processed/Full_Network_output/combined_Network_MR_predicted_age.csv", show_col_types = FALSE)

Network_BrainAGE <- read_csv("/KIMEL/tigrlab/scratch/jshah/brain_age_analysis/data/processed/Full_Network_output/combined_Network_BrainAGE.csv", show_col_types = FALSE)

Network_Adjusted_MR_predicted_age <- read_csv("/KIMEL/tigrlab/scratch/jshah/brain_age_analysis/data/processed/Full_Network_output/combined_Network_Adjusted_MR_predicted_age.csv", show_col_types = FALSE)

#Load demographic file

tay_demo <- read_csv("/KIMEL/tigrlab/scratch/jshah/brain_age_analysis/data/raw/tay_mri_demo_summer2024.csv", show_col_types = FALSE)

#Load data cut file
datacut_ids <- read_csv("/KIMEL/tigrlab/scratch/jshah/brain_age_analysis/data/raw/datacut_ids.csv")

#load euler mean file

euler_data <- read_tsv("/KIMEL/tigrlab/scratch/jshah/brain_age_analysis/data/raw/00_group2_stats_tables/euler.tsv")

```

clean demographic file subject IDs so that they match the output file format to be merged later

```{r}
clean_subject_ids_demo <- function(df, subject_id) {
  df %>%
    mutate(clean_subject_id = str_extract(!!sym(subject_id), "[0-9]{8}$"))
}

TAY_demographics <- clean_subject_ids_demo(tay_demo, "subject_id")

datacut_ids_clean <- clean_subject_ids_demo(datacut_ids, "subject_id")

```
clean subject IDs for the euler_data file

```{r}
clean_subject_ids_tsv <- function(df) {
  colnames(df)[1] <- "clean_subject_id"  # Rename the first column to 'clean_subject_id'
  df %>%
    mutate(clean_subject_id = str_extract(clean_subject_id, "[0-9]{8}$"))  # Extract the last 8 digits of subject IDs
}

euler_data_clean <- clean_subject_ids_tsv(euler_data)

```

turn the subject Ids to numeric because they wont match the output file otherwise

```{r}
TAY_demographics$clean_subject_id <- as.numeric(TAY_demographics$clean_subject_id)
datacut_ids_clean$clean_subject_id <- as.numeric(datacut_ids_clean$clean_subject_id)
euler_data_clean$clean_subject_id <- as.numeric(euler_data_clean$clean_subject_id)

print(TAY_demographics)
print(euler_data_clean)
```

filtered for participant rows that have prodromal psychosis because not all participants have prodromal psychosis listed at T1

```{r}
# Group by subject ID and prioritize rows where symptoms are filled out
# If symptoms are NA, the function will pick the first occurrence that has them filled
TAY_unique_symptoms <- TAY_demographics %>%
  group_by(clean_subject_id) %>%
  arrange(desc(!is.na(prodromal_psychosis))) %>%  # Arrange to prioritize rows where symptoms are NOT NA
  slice(1) %>%  # Keep the first row per subject (the one with symptoms, if available)
  ungroup()

# The result will be a dataframe with unique subject IDs and prioritized symptom data
print(TAY_unique_symptoms)
```

Create a new dataframe with subject IDs from the female template, and age. Then import the PSS/non-PSS from the TAY demographics using the matching subject IDs. 

```{r}
final_analysis_data <- Network_MR_predicted_age %>%
  select(clean_subject_id, age, sex)

final_analysis_data <- final_analysis_data %>%
  left_join(TAY_unique_symptoms %>%
              select(clean_subject_id, prodromal_psychosis),
            by = "clean_subject_id")

```

Add euler's mean

```{r}

final_analysis_data <- final_analysis_data %>%
  left_join(euler_data_clean %>% 
              select(clean_subject_id, mean_euler_bh),  
            by = "clean_subject_id")

```

Add brain age calculator variables

```{r}
final_network_analysis_data<- final_analysis_data %>%
  mutate(adj_pred_error_Vis = Network_Adjusted_BrainAGE$adj_pred_error_Vis, 
         adj_pred_error_SomMot = Network_Adjusted_BrainAGE$adj_pred_error_SomMot,
         adj_pred_error_DorsAttn = Network_Adjusted_BrainAGE$adj_pred_error_DorsAttn,
         adj_pred_error_SalVentAttn = Network_Adjusted_BrainAGE$adj_pred_error_SalVentAttn,
         adj_pred_error_Limbic = Network_Adjusted_BrainAGE$adj_pred_error_Limbic,
         adj_pred_error_Cont = Network_Adjusted_BrainAGE$adj_pred_error_Cont,
         adj_pred_error_Default = Network_Adjusted_BrainAGE$adj_pred_error_Default,
         
         adj_pred_age_Vis = Network_Adjusted_MR_predicted_age$adj_pred_age_Vis, 
         adj_pred_age_SomMot = Network_Adjusted_MR_predicted_age$adj_pred_age_SomMot,
         adj_pred_age_DorsAttn = Network_Adjusted_MR_predicted_age$adj_pred_age_DorsAttn,
         adj_pred_age_SalVentAttn = Network_Adjusted_MR_predicted_age$adj_pred_age_SalVentAttn,
         adj_pred_age_Limbic = Network_Adjusted_MR_predicted_age$adj_pred_age_Limbic,
         adj_pred_age_Cont = Network_Adjusted_MR_predicted_age$adj_pred_age_Cont,
         adj_pred_age_Default = Network_Adjusted_MR_predicted_age$adj_pred_age_Default,
         
         pred_error_Vis = Network_BrainAGE$pred_error_Vis, 
         pred_error_SomMot = Network_BrainAGE$pred_error_SomMot,
         pred_error_DorsAttn = Network_BrainAGE$pred_error_DorsAttn,
         pred_error_SalVentAttn = Network_BrainAGE$pred_error_SalVentAttn,
         pred_error_Limbic = Network_BrainAGE$pred_error_Limbic,
         pred_error_Cont = Network_BrainAGE$pred_error_Cont,
         pred_error_Default = Network_BrainAGE$pred_error_Default,
         
         pred_age_Vis = Network_MR_predicted_age$pred_age_Vis, 
         pred_age_SomMot = Network_MR_predicted_age$pred_age_SomMot,
         pred_age_DorsAttn = Network_MR_predicted_age$pred_age_DorsAttn,
         pred_age_SalVentAttn = Network_MR_predicted_age$pred_age_SalVentAttn,
         pred_age_Limbic = Network_MR_predicted_age$pred_age_Limbic,
         pred_age_Cont = Network_MR_predicted_age$pred_age_Cont,
         pred_age_Default = Network_MR_predicted_age$pred_age_Default,
         )

```

Filter for people who are in the TAY data cut 

```{r}
# Assuming 'sub_ID' is the column to match in both data frames and that `t01` is a logical or character column.
final_network_analysis_data <- final_network_analysis_data %>%
  # Join with datacut_ids_clean to bring in the 't01' column
  left_join(datacut_ids_clean %>% select(clean_subject_id, t01), by = "clean_subject_id")

# View the result
print(final_network_analysis_data)

datacut_filtered_network_data <- final_network_analysis_data %>%
  filter(t01 == TRUE)

# View the result
print(datacut_filtered_network_data)
```

Remove high euler number (cut off above 200)

```{r}
euler_filtered_network_data <- datacut_filtered_network_data %>%
  filter(mean_euler_bh >= -200)

print(euler_filtered_network_data)
```

View all NAs

```{r}
rows_with_na <- euler_filtered_network_data %>%
  filter(if_any(everything(), is.na))
sum(is.na(euler_filtered_network_data$prodromal_psychosis))
sum(is.na(euler_filtered_network_data$adj_pred_age_Default))
print(rows_with_na)
```



Take out NAs from the dataframes

```{r}

cleaned_filtered_network_data <- na.omit(euler_filtered_network_data)

print(cleaned_filtered_network_data)

write.csv(cleaned_filtered_network_data, "/KIMEL/tigrlab/scratch/jshah/brain_age_analysis/data/processed/complete_network_data.csv", row.names = FALSE)

```


Get summary statistics on the dataframe

```{r}

summary(cleaned_filtered_network_data)
table(cleaned_filtered_network_data$sex)

```

```{r}

# Bonferroni-adjusted significance level
bonferroni_alpha <- 0.05 / 7
print(bonferroni_alpha)
```


Run multivariable regression with PSS as the independent variable and adjusted BrainAGE as the dependent variable


```{r}

run_lm_model <- function(dependent_var, independent_vars, data) {
  # Create a formula dynamically
  formula <- as.formula(paste(dependent_var, "~", paste(independent_vars, collapse = " + ")))

  # Run the linear model with the formula and the data explicitly specified
  lm_model <- lm(formula, data = data)

  # Print summary of the model
  return(summary(lm_model))
}

dependent_vars <- c("adj_pred_error_Vis", "adj_pred_error_SomMot", "adj_pred_error_DorsAttn", "adj_pred_error_SalVentAttn", "adj_pred_error_Limbic", "adj_pred_error_Cont", "adj_pred_error_Default", "adj_pred_age_Vis", "adj_pred_age_SomMot", "adj_pred_age_DorsAttn", "adj_pred_age_SalVentAttn", "adj_pred_age_Limbic", "adj_pred_age_Cont", "adj_pred_age_Default", "pred_error_SomMot",      "pred_error_DorsAttn", "pred_error_SalVentAttn", "pred_error_Limbic","pred_error_Cont","pred_error_Default", "pred_age_Vis","pred_age_SomMot","pred_age_DorsAttn", "pred_age_SalVentAttn","pred_age_Limbic",            "pred_age_Cont","pred_age_Default","pred_error_Vis","adj_pred_error_Vis")

for (dep_var in dependent_vars) {
  cat("Running linear model for:", dep_var, "\n")
  lm_summary <- run_lm_model(dep_var, c("prodromal_psychosis", "sex", "mean_euler_bh"), cleaned_filtered_network_data)
  print(lm_summary)  # Print the summary of the model
}

```
```{r}

compute_effect_sizes <- function(var) {
  # Use sym() to allow `cohens_d` to dynamically evaluate the variable name
  cleaned_filtered_network_data %>%
    cohens_d(as.formula(paste(var, "~ prodromal_psychosis")))
}

effect_size_var <- c("adj_pred_error_Vis", "adj_pred_error_SomMot", "adj_pred_error_DorsAttn", "adj_pred_error_SalVentAttn", "adj_pred_error_Limbic", "adj_pred_error_Cont", "adj_pred_error_Default")

effect_sizes_network <- lapply(effect_size_var, compute_effect_sizes)

print(effect_sizes_network)
```

